\bf The forestecology R package for fitting and assessing models of interspecies competitive effects on the growth of trees

### 2021-03-03 paper/forestecology.tex
Done most of introduction

Repeat-censused forest plots offer excellent data to test neighborhood
models of tree competition \citet{allen_permutation_2020}
\citet{canham_neighborhood_2006} \citet{uriarte_spatially_2004}. Here we
describe an R package, \texttt{forestecology}, to do that. This package
implements the methods in \citet{allen_permutation_2020}. It provides: a
convenient way to specify and fit models of tree growth based on
neighborhood competition; a spatial cross validation method to test and
compare model fits \citet{roberts_cross-validation_2017}; and an
ANOVA-like method to assess whether the competitor identity matters in
these models. The model is written to work with ForestGEO plot data
\citet{andersonteixeira_ctfs-forestgeo_2015}, but we envision that it
could easily be modified to work with data from other forest plots,
e.g.~the US Forest Service Forest Inventory and Analysis plots
\citet{smith_forest_2002}. Or more generally to model interactions of
any community of mapped sessile organisms.

The \texttt{forestecology} is designed with ``tidy'' data principles in
mind as \citet{wickham_welcome_2019}.

Given that our data is of geo-spatial nature, we represent our data
using the ``simple features'' \texttt{sf} package class of objects
\citet{pebesma_simple_2018} whereby. While previously the \texttt{sp}
package serves such purposes \citet{pebesma_sp_2005}, the \texttt{sf}
package is designed to interface with the \texttt{tidyverse} suite of
packages.

\hypertarget{competition-model}{%
\subsection{Model for growth of tree}\label{competition-model}}

While there are a litany of models one can consider, in
\citet{allen_permutation_2020} we considered a simple one.

We fit the following linear model to the DBH of each focal tree. Let
\(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
species groups. We modeled the growth in diameter per year \(y_{ij}\)
(in centimeters per year) of the \(i^{th}\) tree of focal species group
\(j\) as a linear model \(f\) of the following covariates
\(\vec{x}_{ij}\)

@@

Link to \url{https://doi.org/10.1371/journal.pone.0229930.s004}

For this linear model's case, there exists a closed form solution as
described here. As such, the \texttt{fit\_bayesian\_model()} function
using matrix algebra to obtain all parameter estimates, rather than
computationally expensive Monte Carlo approximations. The inputs to this
function are a \texttt{focal\_vs\_comp} data frame,
\texttt{prior\_param} a list of priors, and a boolean flag
\texttt{run\_shuffle} on whether or not to run competitor-species
identity permutations which we will demonstrate below on the Michigan
Big Woods data. This function returns the posterior means of all
parameters.

\hypertarget{example}{%
\section{Example}\label{example}}

We demonstrate the \texttt{forestecology} package's functionality on
data from the Smithsonian Conservation Biology Institute (SCBI) large
forest dynamics plot, located at the Smithsonian's National Zoo and
Conservation Biology Institute in Front Royal, VA, USA
\citet{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
the intersection of three of the major physiographic provinces of the
eastern US: the Blue Ridge, Ridge and Valley, and Piedmont provinces and
is adjacent to the northern end of Shenandoah National Park. The forest
type is typical mature secondary eastern mixed deciduous forest, with a
canopy dominated by tulip poplar (\emph{Liriodendron tulipifera}), oaks
(\emph{Quercus} spp.), and hickories (\emph{Carya} spp.), and an
understory composed mainly of spicebush (\emph{Lindera benzoin}),
paw-paw (\emph{Asimina triloba}), American hornbeam (\emph{Carpinus
caroliniana}), and witch hazel (\emph{Hamamelis virginiana})
\citet{bourg_initial_2013}.

@@

The first step in the our analysis sequence is to compute the growth of
trees using data from two censuses. The \texttt{compute\_growth()}
function computes average annual growth assuming census data that
roughly follows ForestGEO standards. Despite such standards, minor
variations will still exist between sites, thereby necessitating some
data wrangling and checking. For example, the SCBI site records all
diameters at breast height (DBH) in millimeters
\citet{bourg_initial_2013}, whereas the Michigan Big Woods site records
them in centimeters \citet{allen_michigan_2020}.

@@

These two data frames are then used as arguments to the
\texttt{compute\_growth()} function, along with the \texttt{id} argument
that specifies the variable that uniquely identifies each tree-stem.
Note furthermore that we discard all resprouts in the later census
(those with \texttt{code\ ==\ R}), since we are only interested in the
diameter growth of surviving, and not resprouted, stems.

@@

The output \texttt{growth\_scbi} is a single data frame of class
\texttt{sf} that includes variables \texttt{growth}, the average annual
growth in DBH (cm y\textsuperscript{-1}) for all stems that were alive
at both time points, and \texttt{geometry}, the \texttt{sf} package's
encoding of geolocations of type
\texttt{\textless{}POINT\textgreater{}}. Furthermore, the variables that
should remain unchanged between censuses appear only once, such as
location variables \texttt{gx} and \texttt{gy}; as well as
species-related variables. Variables that should change between censuses
are suffixed with \texttt{1} and \texttt{2} indicating the earlier and
later censuses, such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.
The data format of other sites may be such that our
\texttt{compute\_growth()} function doesn't work at all. However, in the
end all that matters is that the growth of all stems is saved in a data
frame of class \texttt{sf} and at a minimum the data contains the
variables above. TODO: enumerate what these variables are and say if
they need exact names

@@

The next step in our analysis sequence is to add additional spatial
information to our main \texttt{growth\_scbi} data frame. The first
element of spatial information we add is a ``buffer region'' to the
periphery of the study region. Since some of our model's explanatory
variables are cumulative (such as competitor basal area), we must ensure
that all trees being modeled are not biased to have different neighbor
structures. This is of concern for trees at the boundary of study
regions, for which all neighbors will not be included in the censused
stems. In order to account for such edge effects, only trees that are
not part of this buffer region, i.e.~are part of the interior of the
study region, will have their growth modeled
\citet{waller_applied_2004}.

Our model of interspecific competition relies on a spatial definition of
who the competitor trees are for focal trees of interest: all trees
within a distance \texttt{comp\_dist} of a focal tree are considered its
competitors (assuming the same units as the \texttt{gx} and \texttt{gy}
location variables). In our case we set this value at 7.5m, a value
informed by \citet{canham_neighborhood_2004}
\citet{uriarte_spatially_2004} \citet{canham_neighborhood_2006}. Using
this value along with a manually constructed \texttt{sf} object
representation of the study region's boundary, we apply the
\texttt{add\_buffer\_variable()} to our \texttt{growth\_scbi} data frame
to add a \texttt{buffer} boolean variable: all trees who have
\texttt{buffer} set to \texttt{FALSE} will be our focal trees whose
growth will be modeled, whereas those with \texttt{TRUE} will only be
considered as competitor trees whose growth will not.

@@

The second element of spatial information are blocks corresponding to
folds of a spatial cross-validation algorithm used to estimate
out-of-sample model error. Conventional cross-validation algorithms
assign observations to folds by randomly resampling individual
observations. However, an assumption that the observations are
independent underlies any cross-validation algorithm. In the case of
forest census data, observations exhibit spatial autocorrelation. We
therefore incorporate this spatial dependence into the cross-validation
algorithm with our spatial blocks of trees.
\citet{roberts_cross-validation_2017}
\citet{pohjankukka_estimating_2017}.

In the example below, we first manually define four folds that partition
the study region as an \texttt{sf} object. We then use the output of the
\texttt{spatialBlock()} function from the \texttt{blockCV} package to
associate each tree in \texttt{growth\_scbi} to the correct fold (saved
in the \texttt{foldID} variable) \citet{valavi_blockcv_2019}. \footnote{In
  the Appendix we present an example where the folds themselves are also
  created using the \texttt{spatialBlock()} function given a specified
  \texttt{cv\_block\_size}.}

@@

The next step in our analysis sequence is to identify all focal trees
and their corresponding competitor trees. The
\texttt{create\_focal\_vs\_comp()} functions performs these tasks and
returns a new data frame of type \texttt{sf}. On top of the previous
arguments \texttt{comp\_dist} defining the competition neighborhood and
\texttt{id} indicating which variable uniquely identifies each
tree-stem, this function also requires an \texttt{sf} object
representation of the spatial cross-validation blocks/folds; in our
case, this was manually encoded in the \texttt{blocks\_scbi} in Section
\ref{spatial-information} while in our Appendix we present an example
where this was performed using \texttt{spatialBlock()} from the
\texttt{blockCV} package. We present the resulting data frame below with
the \texttt{foldID} variable omitted for compactness of presentation.

@@

The resulting data frame \texttt{focal\_vs\_comp\_scbi} has 6296 rows,
representing the subset of the 7954 trees in \texttt{growth\_scbi} that
will be considered as focal trees. Two new variables \texttt{focal\_ID}
and \texttt{focal\_sp} relate to tree-stem identification and species
information. Most notably however is a new variable \texttt{comp} which
contains information on all competitor trees for a given focal tree,
saved in \texttt{tidyr} package list-column format
\citet{tidyr_package}. For example, we drill-down on the tree with
\texttt{focal\_ID} 4, which has 20 competitor trees each described by 4
variables as indicated by the fact that \texttt{comp} is a
\texttt{\textless{}tibble\ {[}20\ ×\ 4{]}\textgreater{}}.

@@

Now that we've identified all focal and corresponding competitor trees
and saved this information in a data frame of type
\texttt{focal\_vs\_comp}, the final step in our analysis sequence is to
fit a model for the growth of all focal trees. Currently the
\texttt{forestecology} package can only fit the competition Bayesian
linear regression model outlined in Section \ref{competition-model}
using the \texttt{comp\_bayes\_lm()} function. However, any model
implemented in a function that similarly takes an input data frame of
type \texttt{focal\_vs\_comp} as an argument can be easily swapped in.
For our specific competition Bayesian linear regression model, we also
specify prior distributions on all parameters of interest (here chosen
to be the defaults as specified in \texttt{?comp\_bayes\_lm}).

@@

These plots give the posterior distributions of parameters from Equation
\ref{eq:model}. For many pacakge users they will be of interest because
they give insight into the species-specific competitive interactions.
Setting \texttt{type\ =\ "intercepts"} gives posterior distributions for
\(\beta_{0,j}\) and \texttt{type\ =\ "dbh\_slopes"} for
\(\beta_{dbh,j}\). These give species specific growth independent of
competition. The values of more interest are plotted with
\texttt{type\ =\ \textquotesingle{}competition\textquotesingle{}} which
gives the posterior distribution for \(\lambda_{j,k}\) species-specific
competition coefficients (i.e., the \(\lambda\)-matrix). Negative values
indicate a competitor species which slows the growth of a focal species.
Here, for example, we see that tulip trees (\texttt{litu}) have a strong
negative effect on the growth of conspecifics but relatively little
effect on neighbors of the other two species.

@@

In order to evaluate the effect of competitor species identity, we use
the four steps of our analysis sequence answer along with a permutation
test: Under a null hypothesis where competitor species identity does not
matter, we can permute/shuffle this variable within each focal tree,
compute the RMSE (the test statistic of interest), repeat this process
several times to construct a null distribution of the RMSE, and compare
it to the observed RMSE to assess significance. Going back to our
example in Section \ref{focal-vs-comp} of focal tree with
\texttt{focal\_ID} 4 and its 20 competitors, the permutation test
randomly resamples the \texttt{comp\_sp} variable with replacement,
leaving all other variables intact. The resampling with replacement is
nested within each focal tree in order to preserve the neighborhood
structure of our competition model. To run the permutation test, we use
the same\texttt{comp\_bayes\_lm()} function as in Section
\ref{model-fit-predict}, but with a \texttt{run\_shuffle\ =\ TRUE}
argument.

>>>>

BERT: Dave could you write a few sentences (and add references where
appropriate) starting with the big picture of modeling tree growth,
zooming in to neighhorhood models, and then connecting to the
``\citet{allen_permutation_2020} considers the following model'' I write
below?

Repeat-censused forest plots offer excellent data to test neighborhood
models of the effect of competition on the growth of trees
\citet{canham_neighborhood_2006} \citet{uriarte_spatially_2004}.

\citet{allen_permutation_2020} considers the following model: Let
\(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
species groups. We model the average annual growth in diameter
\(y_{ij}\) (in centimeters per year) of the \(i^{th}\) tree of focal
species group \(j\) as a linear model \(f\) of the covariates
\(\vec{x}_{ij}\)

@@

where \(\beta_{0,j}\) is the diameter-independent growth rate for group
\(j\); \(\text{dbh}_{ij}\) is the diameter at breast height (in
centimeters) of the focal tree at the earlier census;
\(\beta_{\text{dbh},j}\) is the amount of the growth rate changed
depending on diameter for group \(j\); \(\text{BA}_{ijk}\) is the sum of
the basal area of all trees of competitor species group \(k\);
\(\lambda_{jk}\) is the change in growth for individuals of group \(j\)
from nearby competitors of group \(k\); and \(\epsilon_{ij}\) is a
random error term distributed \(\text{Normal}(0, \sigma^2)\). They
estimate all parameters via Bayesian linear regression while exploiting
Normal/Inverse Gamma conjugacy to derive closed-form solutions to all
posterior distributions via linear algebra\footnote{See S1 Appendix of
  \citet{allen_permutation_2020}, available at
  \url{https://doi.org/10.1371/journal.pone.0229930.s004}}. These
closed-form solutions for the posterior distributions are in contrast to
approximations of all posteriors via computationally expensive Markov
Chain Monte Carlo algorithms.

In order to evaluate whether competitor species identity matters,
\citet{allen_permutation_2020} run a permutation test where under the
null hypothesis the species identity of all competitors of a focal tree
can be permuted/shuffled:

\begin{eqnarray}
\label{eq:permutation-hypothesis-test}
&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K\\
\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different}
\end{eqnarray}

where the null hypothesis \(H_0\) reflects a hypothesis of no species
grouping-specific effects of competition while the alternative
hypothesis \(H_A\) reflects a hypothesis of species grouping-specific
effects of competition. Furthermore, in order to account for the spatial
autocorrelation inherent to forest data in their estimates of
out-of-sample model error, \citet{allen_permutation_2020} use spatial
cross-validation. Estimates of model error that do not account for this
spatial dependency tend to underestiamte the true model error
\citep{roberts_cross-validation_2017}.

We introduce the \texttt{forestecology} R package providing methods and
data for forest ecology model fitting and assessment, available on CRAN
(\url{https://cran.r-project.org/web/packages/forestecology/index.html})
with the corresponding source code available on GitHub
(\url{https://github.com/rudeboybert/forestecology}). The package
implements all aspects of the model in Equation \ref{eq:model}: model
fitting and generating fitted/predicted values, evaluating the effect of
competitor species identity using permutation tests, and evaluating
model performance using spatial cross-validation.

The package designed with ``tidy'' design principles in mind
\citep{wickham_welcome_2019}. Much like many of the \texttt{tidyverse}
component packages, \texttt{forestecology} is designed with verb-named
functions that can be modularly composed in sequence using the pipe
\texttt{\%\textgreater{}\%} operator \citep{bache_pipe_2020}. As we
articulate in Section \ref{casestudy}, these functions delineate the key
steps in our analysis sequence. Furthermore, the inputs and outputs of
nearly all of our functions use the same ``simple features for R'' data
structures as implemented in the \texttt{sf} package for standardized
support for spatial vector data \citep{pebesma_simple_2018}. The
\texttt{sf} package is a \texttt{tidyverse}-friendly evolution of the
\texttt{sp} package of classes and methods for spatial data in R
\citep{pebesma_sp_2005}. As such, wrangling and visualization spatial
data such as ours becomes much easier.

Currently the package only implements the Bayesian linear regression
model of tree growth based on neighborhood competition detailed in
Equation \ref{eq:model}. As we demonstrate in Section
\ref{model-fit-predict} however, the fitting of this model is
self-contained in a single function \texttt{comp\_bayes\_lm()}. This
function returns an object of S3 class type \texttt{comp\_bayes\_lm}
with generic methods implemented for \texttt{print()} to inspect the
output, \texttt{predict()} to generate fitted/predicted values, and
\texttt{ggplot2::autoplot()} to visualize all results. Therefore the
package can be modularly extended to fit other models as long as they
are coded into a function similar type as \texttt{comp\_bayes\_lm()} as
has equivalent generic methods implemented.

We present a case-study of the \texttt{forestecology} package's use on
data from the Smithsonian Conservation Biology Institute's (SCBI) large
forest dynamics plot in Front Royal, Virginia, USA in Section
\ref{casestudy}, which is part of the ForestGEO global network of
research sites
\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}.
The package is designed with ForestGEO plot data in mind, but we
envision that it could easily be modified to work with data from other
forest plots, e.g.~the US Forest Service Forest Inventory and Analysis
plots or more generally to model interactions of any community of mapped
sessile organisms \citep{smith_forest_2002}.

\hypertarget{casestudy}{%
\section{forestecology workflow: a case study}\label{casestudy}}

We demonstrate the \texttt{forestecology} package's functionality on
data from the Smithsonian Conservation Biology Institute (SCBI) large
forest dynamics plot, located at the Smithsonian's National Zoo and
Conservation Biology Institute in Front Royal, VA, USA
\citep{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
the intersection of three of the major physiographic provinces of the
eastern US---the Blue Ridge, Ridge and Valley, and Piedmont
provinces---and is adjacent to the northern end of Shenandoah National
Park. The forest type is typical mature secondary eastern mixed
deciduous forest, with a canopy dominated by tulip poplar
(\emph{Liriodendron tulipifera}), oaks (\emph{Quercus} spp.), and
hickories (\emph{Carya} spp.), and an understory composed mainly of
spicebush (\emph{Lindera benzoin}), paw-paw (\emph{Asimina triloba}),
American hornbeam (\emph{Carpinus caroliniana}), and witch hazel
(\emph{Hamamelis virginiana}) \citep{bourg_initial_2013}.

@@

The first step in the our analysis sequence is to compute the growth of
trees using data from two censuses. The \texttt{compute\_growth()}
function computes average annual growth assuming census data that
roughly follows ForestGEO standards. Despite such standards, minor
variations will still exist between sites, thereby necessitating some
data wrangling and checking. For example, the SCBI site records all
diameters at breast height (DBH) in millimeters
\citep{bourg_initial_2013}, whereas the Michigan Big Woods site records
them in centimeters \citep{allen_michigan_2020}.

@@

These two data frames are then supplied as arguments to the
\texttt{compute\_growth()} function, along with the \texttt{id} argument
that specifies the variable that uniquely identifies each tree-stem.
Note furthermore that we discard all resprouts in the later census
(those with \texttt{code\ ==\ R}), since we are only interested in the
diameter growth of surviving, and not resprouted, stems.

@@

The output \texttt{growth\_scbi} is a single data frame of class
\texttt{sf} that includes variables \texttt{growth}, the average annual
growth in DBH (cm y\textsuperscript{-1}) for all stems that were alive
at both time points, and \texttt{geometry}, the \texttt{sf} package's
encoding of geolocations of type
\texttt{\textless{}POINT\textgreater{}}. Furthermore, the variables that
should remain unchanged between censuses appear only once, such as
location variables \texttt{gx} and \texttt{gy}; as well as
species-related variables. Variables that should change between censuses
are suffixed with \texttt{1} and \texttt{2} indicating the earlier and
later censuses, such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.
Site data that does not align with this convention will need to be
transformed for use with the \texttt{compute\_growth()} function.
However, in the end, all that matters is that the growth of all stems is
saved in a data frame of class \texttt{sf} and, at a minimum, the data
contains the variables above. TODO: enumerate what these variables are
and say if they need exact names

@@

The next step in our analysis sequence is to add additional spatial
information to our main \texttt{growth\_scbi} data frame. The first
element of spatial information we add is a ``buffer region'' to the
periphery of the study region. Since some of our model's explanatory
variables are cumulative (such as competitor basal area), we must ensure
that all trees being modeled are not biased to have different neighbor
structures. This is of concern for trees at the boundary of study
regions, for which all neighbors will not be included in the censused
stems. In order to account for such edge effects, only trees that are
not part of this buffer region, i.e.~are part of the interior of the
study region, will have their growth modeled
\citep{waller_applied_2004}.

Our model of interspecific competition relies on a spatial definition of
who the competitor trees are for focal trees of interest: all trees
within a distance \texttt{comp\_dist} of a focal tree are considered its
competitors (assuming the same units as the \texttt{gx} and \texttt{gy}
location variables). In our case we set this value at 7.5m, a value
informed by other studies \citep[\citet{uriarte_spatially_2004},
\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. Using this
value along with a manually constructed \texttt{sf} object
representation of the study region's boundary via its vertices, we apply
the \texttt{add\_buffer\_variable()} to our \texttt{growth\_scbi} data
frame to add a \texttt{buffer} boolean variable: all trees who have
\texttt{buffer} set to \texttt{FALSE} will be our focal trees whose
growth will be modeled, whereas those with \texttt{TRUE} will only be
considered as competitor trees whose growth will not.

@@

The second element of spatial information are blocks corresponding to
folds of a spatial cross-validation algorithm used to estimate
out-of-sample model error. Conventional cross-validation algorithms
assign observations to folds by randomly resampling individual
observations. However, many of these algorithms assume that the
observations are independent of each other. In the case of forest census
data, observations exhibit spatial autocorrelation. We therefore
incorporate this spatial dependence into the cross-validation algorithm
with our spatial blocks of trees \citep[
\citet{pohjankukka_estimating_2017}]{roberts_cross-validation_2017} In
the example below, we first manually define four folds that partition
the study region as an \texttt{sf} object. We then use the output of the
\texttt{spatialBlock()} function from the \texttt{blockCV} package to
associate each tree in \texttt{growth\_scbi} to the correct fold (saved
in the \texttt{foldID} variable) \citep{valavi_blockcv_2019}. \footnote{In
  the Appendix we present an example where the folds themselves are also
  created using the \texttt{spatialBlock()} function given a specified
  \texttt{cv\_block\_size}.}

@@

The next step in our analysis sequence is to identify all focal trees
and their corresponding competitor trees. The
\texttt{create\_focal\_vs\_comp()} functions performs these tasks and
returns a new data frame of type \texttt{sf}. On top of the previous
arguments \texttt{comp\_dist} defining the competition neighborhood and
\texttt{id} indicating which variable uniquely identifies each
tree-stem, this function also requires an \texttt{sf} object
representation of the spatial cross-validation blocks/folds. In this
example, the blocks were manually encoded in \texttt{blocks\_scbi} by
specifying it's vertices in Section \ref{spatial-information}\footnote{We
  present an alternative method for defining spatial cross-validation
  blocks is using the \texttt{spatialBlock()} function from the
  \texttt{blockCV} package in the Appendix}. We present the resulting
data frame below with the \texttt{foldID} variable omitted for
compactness of presentation.

@@

The resulting data frame \texttt{focal\_vs\_comp\_scbi} has 6296 rows,
representing the subset of the 7954 trees in \texttt{growth\_scbi} that
will be considered as focal trees. Two new variables \texttt{focal\_ID}
and \texttt{focal\_sp} relate to tree-stem identification and species
information. Most notably however is a new variable \texttt{comp} which
contains information on all competitor trees for a given focal tree,
saved in \texttt{tidyr} package list-column format
\citep{tidyr_package}. For example, we drill-down on the tree with
\texttt{focal\_ID} 4, which has 20 competitor trees each described by 4
variables as indicated by the fact that \texttt{comp} is a
\texttt{tibble\ {[}20\ ×\ 4{]}}.

@@

Now that we've identified all focal and corresponding competitor trees
and saved this information in a data frame of type
\texttt{focal\_vs\_comp}, the final step in our analysis sequence is to
fit a model for the growth of all focal trees. Currently the
\texttt{forestecology} package can only fit the competition Bayesian
linear regression model outlined in Section \ref{competition-model}
using the \texttt{comp\_bayes\_lm()} function. However, any model
implemented in a function that similarly takes an input data frame of
type \texttt{focal\_vs\_comp} as an argument can also be used. For our
specific competition Bayesian linear regression model, we also specify
prior distributions on all parameters of interest (here chosen to be the
defaults as specified in \texttt{?comp\_bayes\_lm}).

@@

These plots give the posterior distributions of parameters from Equation
\ref{eq:model}. For many package users they will be of interest because
they give insight into the species-specific competitive interactions.
Setting \texttt{type\ =\ "intercepts"} gives posterior distributions for
\(\beta_{0,j}\) and \texttt{type\ =\ "dbh\_slopes"} for
\(\beta_{dbh,j}\). These give species specific growth independent of
competition. The values of more interest are plotted with
\texttt{type\ =\ \textquotesingle{}competition\textquotesingle{}} which
gives the posterior distribution for \(\lambda_{j,k}\) species-specific
competition coefficients (i.e., the \(\lambda\)-matrix). Negative values
indicate a competitor species which slows the growth of a focal species.
Here, for example, we see that tulip trees (\texttt{litu}) have a strong
negative effect on the growth of conspecifics but relatively little
effect on neighbors of the other two species.

@@

In order to evaluate the effect of competitor species identity, we use
the four steps of our analysis sequence answer along with a permutation
test: Under a null hypothesis where competitor species identity does not
matter, we can permute/shuffle this variable within each focal tree,
compute the RMSE (the test statistic of interest), repeat this process
several times to construct a null distribution of the RMSE, and compare
it to the observed RMSE to assess significance. Going back to our
example in Section \ref{focal-vs-comp} of focal tree with
\texttt{focal\_ID} 4 and its 20 competitors, the permutation test
randomly resamples the \texttt{comp\_sp} variable with replacement,
leaving all other variables intact. The resampling with replacement is
nested within each focal tree in order to preserve the neighborhood
structure of our competition model. To run the permutation test, we use
the same \texttt{comp\_bayes\_lm()} function as in Section
\ref{model-fit-predict}, but with a \texttt{run\_shuffle\ =\ TRUE}
argument.
